Approximate State Abstraction

### 2016-01-29 paper/paper.tex
Mostly fixed Q* proof

% --- Notation Commands ---
\newcommand{\ep}{\widetilde \phi}
\newcommand\defn[1]{{\bf Definition:} #1}

@@

% Subsection:  Similarity of Value = Similar Behavior
\subsection{Similarity of Value Yields Similar Behavior}

{\bf Lemma 1:}
If the value of a state in the ground \ac{MDP} is within $\kappa \in \mathbb{R}$ of the value of the corresponding abstract state, then the value of the ground state under the optimal policy in the abstract \ac{MDP} is within $\kappa$ of the value of the ground state:
\begin{multline*}
| V_G(s) - V_A(\phi (s)) | \leq \kappa \rightarrow
\left| V_G(s) - V_G^{\pi_A^*} \right| \leq \kappa
\end{multline*}

\textbf{Proof:}

@@

% Q* Lemma
Recall that $Q^*$ is a function on $\mathcal{S}_G \times \mathcal{A}$ that satisfies \enote{$Q^*$ definition}. Inspired by the $Q^*$ irrelevance abstraction, we consider the approximate case. By Equation~\ref{eq:phi_f}, $\ep_{Q^*}$ is a type of abstraction that, for a fixed $\epsilon$, satisfies:
\begin{equation}
\ep_{Q^*}(s_1) = \ep_{Q^*}(s_2) \rightarrow \forall_a \left|Q^*(s_1, a) - Q^*(s_2, a)\right| \leq \epsilon
\end{equation}

{\bf Lemma 2:} When $S_A$ is created using a function of the $\ep_{Q^*}$ type:
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{\epsilon}{1-\gamma}
\end{equation}

\textbf{Proof:}
Consider a temporally heterogeneous \ac{MDP}, $M^T$, parameterized by integer $T$, such that for the first $T$ time steps the reward function, transition dynamics and state space are those of the abstract MDP, $M_A$, and after $T$ time steps the reward function, transition dynamics and state spaces are those of the ground MDP, $M_G$. Note that $M^0 = M_G$ and $M^\infty = M_A$.

@@

@@

And similarly:
\begin{equation}
\mathcal{T}_T = \begin{cases}
\mathcal{T}_A& T > 0 \\
\mathcal{T}_G& \text{otherwise} \\
\end{cases}
\end{equation}

We proceed by induction on $T$ to show that:
\begin{equation}
\forall_T |V^{\pi_G^*}(s) - V_T(s)| \leq \sum_{t=0}^{T+1} \epsilon \gamma^t
\end{equation}

@@

When T = 0, $V^{\pi_G^*}(s) = V_0(s)$, so it trivially follows that:
\begin{equation}
|V^{\pi_G^*}(s) - V_0(s)| \leq \epsilon = \sum_{t=0}^{0} \epsilon \gamma^t
\end{equation}

% Inductive Case
\textit{Inductive Case:}

We assume as our inductive hypothesis that:
\begin{equation}
\forall_s\ V_{T-1}(s) \leq \frac{\epsilon}{1-\gamma} + V^{\pi_G^*}(s)
\end{equation}
\begin{align*}
&\text{By definition of $V_{T}(s)$:}\\
&V_{T}(s) = \max_a \left[R_{T}(s,a) + \gamma \sum_s' T_{T}(s,a,s')V_{T-1}(s')\right] \\
&\leq \max_a \left[R_{T}(s,a) + \gamma \sum_s' T_{T}(s,a,s')\left(V_0(s) + \sum_{t=0}^{T-1}\epsilon\gamma^t \right) \right] \\
&= \gamma\left(\sum_{t=0}^{T-1}\epsilon\gamma^{t}\right) + \\ &\max_a \left[R_{T}(s,a) + \gamma \sum_s' T_{T}(s,a,s')V_0(s') \right] \\
\end{align*}
\begin{align*}
&= \gamma\left(\sum_{t=0}^{T-1}\epsilon\gamma^t\right) + \\ &\max_a \left[\sum_{g \in s} R_{G}(g,a)\omega(g) + \gamma \sum_{s'} \sum_{g \in s}T_G(g,a,s')\omega(g)V_0(s') \right] \\
&= \gamma\left(\sum_{t=0}^{T-1}\epsilon\gamma^t\right) + \epsilon + \left(\max_a Q_G(s,a)\right) \\
&= \sum_{t=0}^{T} \epsilon \gamma^{t} + \max_a Q^{\pi_G^*}(s,a) \\
&= \frac{\epsilon}{1-\gamma} + V^{\pi_G^*}(s) \\
\end{align*}

Therefore, for all $T$:
\begin{align*}
V_{T}(s) &\leq \frac{\epsilon}{1-\gamma} + V^{\pi_G^*}(s) \\
|V^{\pi_G^*}(s) - V_{T}(s)| &\leq \frac{\epsilon}{1-\gamma}
\end{align*}

Note that as $T \rightarrow \infty$, $V_T$ becomes $V^{\pi_A*}$. Therefore:

\begin{equation}
|V^{\pi_G^*}(s) - V^{\pi_A^*}(s)| \leq \frac{\epsilon}{1-\gamma}
\end{equation}

% Subsection: Model Similarity
\subsection{Model Similarity: $\ep_{model}$}

@@

\begin{align*}
B_{s} &= \max_{s,a}|Q'(s,a)-Q(s,a)|\\
%B_{s} &= \max_{s,a}|R'(s,a) - R(s,a) + \gamma \sum_{s'}\left[T'(s,a,s')\max_{a'}Q'(s',a')-T(s,a,s')\max_{a'}Q(s', a')\right]|\\
&\forall s' \text{ WLOG } T'(s,a,s')\max_{a'}Q'(s',a') \geq T(s,a,s')\max_{a'}Q(s',a')\\
\end{align*}
Applying max is a non-expansion:
\begin{multline}
=\max_{s,a}(|R'(s,a)-R(s,a)| + \gamma \sum_{s'} \\
\max_{a'} [(T(s,a,s')+T'(s,a,s')- \\ T(s,a,s'))Q'(s', a')-T(s,a,s')Q(s',a')])
\end{multline}
%&\leq \max_{s,a}|R'_s-R_s|+\max_{s,a}\gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')+\epsilon_{\mathcal{T}})Q'(s',a') - T(s,a,s')Q(s',a')\right]\\
%&\leq \epsilon_{\mathcal{R}}+\max_{s,a}\gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')Q'(s',a')+\epsilon_{\mathcal{T}}Q'(s',a') - T(s,a,s')Q(s',a')\right]\\
%&\leq \epsilon_{\mathcal{R}} + \max_{s,a} \gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')(Q'(s',a')-Q(s',a'))+\epsilon_{\mathcal{T}}Q'(s',a') \right]\\
\begin{align*}
&\leq \epsilon_{\mathcal{R}}+\gamma |\mathcal{S}|\epsilon_{\mathcal{T}}Q'_{\text{max}} + \gamma \sum_{s'}\left[ T(s,a,s') B_s\right]\\
&\leq \epsilon_{\mathcal{R}} + \gamma B_{s} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}}Q'_{max}
\end{align*}

@@

% Subsection: Multinomial over Optimal Q
\subsection{Multinomial over Optimal Q: $\ep_{mult}$}

>>>>

% --- Notation Commands ---
\newcommand{\ep}{\widetilde \phi}
\newcommand{\epQ}{\ep_{Q^*}}
\newcommand\defn[1]{{\bf Definition:} #1}

@@

@@

% Q* Lemma
Recall that $Q^*$ is a function on $\mathcal{S}_G \times \mathcal{A}$ that satisfies \enote{$Q^*$ definition}. Inspired by the $Q^*$ irrelevance abstraction, we consider the approximate case. By Equation~\ref{eq:phi_f}, $\ep_{Q^*}$ is a type of abstraction that, for a fixed $\epsilon$, satisfies:

\begin{equation}
\ep_{Q^*}(s_1) = \ep_{Q^*}(s_2) \rightarrow \forall_a \left|Q^*(s_1, a) - Q^*(s_2, a)\right| \leq \epsilon
\end{equation}

{\bf Lemma 1:} When $S_A$ is created using a function of the $\ep_{Q^*}$ type:
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{\epsilon}{(1-\gamma)^2}
\end{equation}

\textbf{Proof:}
We begin by showing that Q values in the abstract \ac{MDP} closely resemble those in the ground \ac{MDP}.

Consider a temporally heterogeneous \ac{MDP}, $M^T$, parameterized by integer $T$, such that for the first $T$ time steps the reward function, transition dynamics and state space are those of the abstract MDP, $M_A$, and after $T$ time steps the reward function, transition dynamics and state spaces are those of the ground MDP, $M_G$. Note that $M^0 = M_G$ and $M^\infty = M_A$.

@@

Similarly, the Q-value of state $s$ in $M^T$ for action $a$ is defined by:
\begin{equation}
Q(s, a) = \mathcal{R}_T(s,a) + \gamma \sum_{s'} \mathcal{T}_T(s,a,s') \max_{a'} Q(s',a')
\end{equation}

@@

And:
\begin{equation}
\mathcal{T}_T = \begin{cases}
\mathcal{T}_A& T > 0 \\
\mathcal{T}_G& \text{otherwise} \\
\end{cases}
\end{equation}

\enote{Add something about weirdness of state spaces changing at $T=1$}

We proceed by induction on $T$ to show that:
\begin{equation}
\forall_{T, g \in \mathcal{S}_G, a} |Q_G(g, a) - Q_T(\epQ(g), a)| \leq \sum_{t=0}^{T+1} \epsilon \gamma^t
\end{equation}

@@

\enote{Fix this -- state spaces are the same for this case}

When T = 0, $Q^{\pi_G^*} = Q_0$, so it trivially follows that:
\begin{equation}
\forall_{g\in \mathcal{S_G}, a} |Q^{\pi_G^*}(g, a) - Q_T(\ep_{Q^*}(g), a)| \leq \sum_{t=0}^{0} \epsilon \gamma^t
\end{equation}

\enote{Add base case for $T=1$ since our value function is not well defined unless $T > 1$.}

% Inductive Case
\textit{Inductive Case:} \enote {For $T>1$}

We assume as our inductive hypothesis that:
\begin{equation}
\forall_{g \in \mathcal{S}_G, a} |Q^{\pi_G^*}(g, a) - Q_{T-1}(\epQ(g), a)| \leq \sum_{t=0}^{T-1} \epsilon \gamma^t\\
\end{equation}

Consider a fixed but arbitrary state, $s_G \in \mathcal{S}_G$, and fixed but arbitrary action $a$.

We denote $\epQ(s_G)$ as $s_A$.
\begin{align*}
&\text{By definition of $Q_{T}(s_A, a)$:}\\
&Q_T(s_A, a) = \mathcal{R}_T(s_A,a) + \gamma \sum_{{s_A}' \in M_T} \mathcal{T}_T(s_A,a,{s_A}') \max_{a'} Q_{T-1}({s_A}',a')\\
&=\sum_{g}w(g)\left[ R_0(g,a) + \gamma \sum_{g'} T_0(g,a,g') \max_{a'} Q_{T-1}(g', a')      \right]\\
&\text{Where $g$ ranges over states such that $\epQ(g) = S_A$ and $g' \in \mathcal{S}_G$}\\
&\text{Applying our inductive hypothesis yields}\\
&\leq \sum_{g}w(g)\left[ R_0(g,a) + \gamma \sum_{g'} T_0(g,a,g') \max_{a'}(Q_0(g', a') + \sum_{t=0}^{T-1} \epsilon \gamma^t)      \right]\\
&\leq \gamma\sum_{t=0}^{T-1} \epsilon \gamma^t + \sum_{g}w(g)\left[ R_0(g,a) + \gamma \sum_{g'} T_0(g,a,g') \max_{a'}Q_0      \right]\\
&\leq \gamma\sum_{t=0}^{T-1} \epsilon \gamma^t + \sum_{g}\left[ w(g)\ Q_0(g,a)\right]\\
&\text{Since all aggregated states have Q-values within $\epsilon$ of one another:}\\
&\leq \gamma\sum_{t=0}^{T-1} \epsilon \gamma^t + \epsilon + Q_0(s_G, a)\\
&\leq \gamma\sum_{t=0}^{T} \epsilon \gamma^t + Q_0(s_G, a)\\
&\text{Thus,}\\
&\left| Q_{T}(s_A, a) - Q_0(s_G,a) \right| \leq \gamma\sum_{t=0}^{T}\epsilon \gamma^t
\end{align*}

Since $s_G$ and $s_A$ are arbitrary, we conclude
\begin{equation}
\forall_{T, g \in \mathcal{S}_G, a} |Q^{\pi_G^*}(g, a) - Q_T(\epQ(g), a)| \leq \sum_{t=0}^{T} \epsilon \gamma^t
\end{equation}

%\begin{align*}
%&= \gamma\left(\sum_{t=0}^{T-1}\epsilon\gamma^t\right) + \\ &\max_a \left[\sum_{g \in s} R_{G}(g,a)\omega(g) + \gamma \sum_{s'} \sum_{g \in s}T_G(g,a,s')\omega(g)V_0(s') \right] \\
%&= \gamma\left(\sum_{t=0}^{T-1}\epsilon\gamma^t\right) + \epsilon + \left(\max_a Q_G(s,a)\right) \\
%&= \sum_{t=0}^{T} \epsilon \gamma^{t} + \max_a Q^{\pi_G^*}(s,a) \\
%&= \frac{\epsilon}{1-\gamma} + V^{\pi_G^*}(s) \\
%\end{align*}

As $T \rightarrow \infty$, $Q_T$ becomes $Q_A$ and $\sum_{t=0}^{T} \epsilon \gamma^t$ becomes $\frac{\epsilon}{1-\gamma}$. Therefore:

\begin{equation}
\label{eq:Q*HalfResult}
\forall_{g \in \mathcal{S}_G, a} |Q_G(g, a) - Q_A(\epQ(g), a)| \leq \frac{\epsilon}{1-\gamma}
\end{equation}

max

Having proved that the Q values in the abstract MDP formed by abstractions of the $\epQ$ type closely resemble those in the ground \ac{MDP}, we now use this result to demonstrate that optimal behavior in the abstract \ac{MDP} is nearly optimal in the ground \ac{MDP}.

Consider the following policy on states in $\mathcal{S}_G$:
\begin{equation}
\pi_{A,t}(s)=
\begin{cases}
\pi_G^*(s), \text{if $t<=0$}\\
\pi_A^*(\ep(s)), \text{if $t > 0$}\\
\end{cases}
\end{equation}

\begin{align*}
\text{Let }Q^{\pi_{A,t}}(g) &= \\
&R_G(s, \pi_{A,t}(g)) + \gamma \sum_{g'}\mathcal{T}_G(g, a, g')\max_{a'}Q^{\pi_{A,t-1}}(g', a')
\end{align*}

where $g, g' \in \mathcal{S}_G$.

We now show by induction on $t$ that
\begin{equation}
\forall_{t, g \in \mathcal{S}_g, a} Q^{\pi_{A,t}}(g, a) \leq Q^{\pi^*}(g,a) + \sum_{i=0}^{t}\gamma^i \frac{\epsilon}{1-\gamma}
\end{equation}

\textit{Base case: $t=0$}

By definition when $t=0$, $Q^{\pi_{A,t}} = Q^{\pi^*}$ so it trivially follows that
\begin{equation}
\forall_{t, g \in \mathcal{S}_g, a} Q^{\pi_{A,0}}(g, a) \leq Q^{\pi^*}(g,a) + \frac{\epsilon}{1-\gamma}
\end{equation}

\textit{Inductive case: $t > 0$}

We assume for our inductive hypothesis that
\begin{equation}
\forall_{t, g \in \mathcal{S}_g, a} Q^{\pi_{A,t-1}}(g, a) \leq Q^{\pi^*}(g,a) + \sum_{i=0}^{t-1}\gamma^i \frac{\epsilon}{1-\gamma}
\end{equation}

By definition 
\begin{align*}
&Q^{\pi_{A,t}}(g, a) = R_G(s, \pi_{A,t}(g)) + \gamma \sum_{g'}\mathcal{T}_G(g, a, g')\max_{a'}Q^{\pi_{A,t-1}}(g', a')\\
&\text{Applying our inductive hypothesis yields}\\
&Q^{\pi_{A,t}}(g, a) \leq\\
&R_G(s, \pi_{A,t}(g, a)) + \gamma \sum_{g'}\mathcal{T}_G(g, a, g')\max_{a'}\left(Q^{\pi^*}(g',a') + \sum_{i=0}^{t-1}\gamma^i \frac{\epsilon}{1-\gamma} \right)\\
&\leq \gamma\sum_{i=0}^{t-1}\gamma^i \frac{\epsilon}{1-\gamma} + R_G(s, \pi_{A,t}(g)) + \gamma \sum_{g'}\mathcal{T}_G(g, a, g')\max_{a'}Q^{\pi^*}(g',a')\\
&\leq \gamma\sum_{i=0}^{t-1}\gamma^i \frac{\epsilon}{1-\gamma} + Q^{\pi^*}(g, \pi_{A,t} (g))\\
&\text{Applying Equation \ref{} yields}\\
&\leq \gamma\sum_{i=0}^{t-1}\gamma^i \frac{\epsilon}{1-\gamma} + Q^{\pi^*}(g, a)\\
\end{align*}
% Subsection: Model Similarity
\subsection{Model Similarity: $\ep_{model}$}

@@

\begin{align*}
B_{s} &= \max_{s,a}|Q'(s,a)-Q(s,a)|\\
%B_{s} &= \max_{s,a}|R'(s,a) - R(s,a) + \gamma \sum_{s'}\left[T'(s,a,s')\max_{a'}Q'(s',a')-T(s,a,s')\max_{a'}Q(s', a')\right]|\\
&\forall s' \text{ WLOG } T'(s,a,s')\max_{a'}Q'(s',a') \geq T(s,a,s')\max_{a'}Q(s',a')\\
\end{align*}
Applying max is a non-expansion \enote{Probably want a citation and or explanation for this}:
\begin{multline}
=\max_{s,a}(|R'(s,a)-R(s,a)| + \gamma \sum_{s'} \\
\max_{a'} [(T(s,a,s')+T'(s,a,s')- \\ T(s,a,s'))Q'(s', a')-T(s,a,s')Q(s',a')])
\end{multline}
%&\leq \max_{s,a}|R'_s-R_s|+\max_{s,a}\gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')+\epsilon_{\mathcal{T}})Q'(s',a') - T(s,a,s')Q(s',a')\right]\\
%&\leq \epsilon_{\mathcal{R}}+\max_{s,a}\gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')Q'(s',a')+\epsilon_{\mathcal{T}}Q'(s',a') - T(s,a,s')Q(s',a')\right]\\
%&\leq \epsilon_{\mathcal{R}} + \max_{s,a} \gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')(Q'(s',a')-Q(s',a'))+\epsilon_{\mathcal{T}}Q'(s',a') \right]\\
\begin{align*}
&\leq \epsilon_{\mathcal{R}}+\gamma |\mathcal{S}|\epsilon_{\mathcal{T}}Q'_{\text{max}} + \gamma \sum_{s'}\left[ T(s,a,s') B_s\right]\\
&\leq \epsilon_{\mathcal{R}} + \gamma B_{s} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}}Q'_{max}
\end{align*}

@@

% Subsection: Multinomial over Optimal Q
\enote{If not here, we should have some motivation for multinomial (and boltz)}
\subsection{Multinomial over Optimal Q: $\ep_{mult}$}
